State of the Art in Concurrent Priority Queues \\ Student Paper for Seminar in Algorithms 2013W \\ Technical University of Vienna

### 2014-03-17 paper/paper.tex
Use \citeauthor in section titles
Finish up the Linden section

\subsection{Shavit \& Lotan} \label{sec:shavit}

@@

\subsection{Sundell \& Tsigas} \label{sec:sundell}

@@

\subsection{Linden \& Jonsson} \label{sec:linden}

@@

As in Section \ref{sec:sundell}, deletion flags are packed into each node's pointers. However,
% Deletion flag only on next[0]!

%  ``Entirely limited by logical deletion in DeleteMin''

 % Mention the BoundOffset parameter (p17), and benching using Fraser-based impls.

@@

\subsection{Wimmer} \label{sec:wimmer}

>>>>

\subsection{\citeauthor{shavit2000skiplist}} \label{sec:shavit}

@@

\subsection{\citeauthor{sundell2003fast}} \label{sec:sundell}

@@

\subsection{\citeauthor{linden2013skiplist}} \label{sec:linden}

@@

In this implementation, \lstinline|DeleteMin| operations only perform logical deletion by setting
the deletion flag with a single \ac{FAO} call; nodes are only deleted physically once a certain
threshold of logically deleted nodes is reached.

This mechanism requires a new invariant, in that the set of all logically deleted nodes must always
form a prefix of the SkipList. Recall that in the \citeauthor{sundell2003fast} queue, deletion flags
for node \lstinline|n| were packed into \lstinline|n.next| pointers, preventing insertion of new
nodes \emph{after} deleted nodes. This implementation instead places the deletion flag into the
lowest level \lstinline|next| pointer of the previous node, preventing inserations \emph{before}
logically deleted nodes.

Once the prefix of logically deleted nodes reaches a specified length \lstinline|BoundOffset|,
the first thread to observe this fact within \lstinline|DeleteMin| performs the actual physical
deletions by updating \lstinline|slist.head[0]| to point at the last logically deleted node with a
single \ac{FAO} operation. The remaining \lstinline|slist.head| pointers are then updated, and
all physically deleted nodes are marked for recycling.

Since at any time, the data structure contains a prefix of logically deleted nodes, all \lstinline|DeleteMin|
operations must traverse this sequence before reaching a non-deleted node. In general, reads of nonmodified
memory locations are very cheap; however, benchmarks in \cite{linden2013skiplist} have shown that
after a certain point, the effort spent in long read sequences significantly outweighs the reduced
number of \ac{CAS} calls. It is therefore crucial to choose \lstinline|BoundOffset| carefully, with the
authors recommending a prefix length bound of 128 for 32 threads.

The actual \lstinline|DeleteMin| and \lstinline|Insert| implementations are surprisingly simple.
Deletions simply traverse the list until the first node for which \lstinline|(ptr, d) = FAO((node.next[0], d), 1)|
returns a previously unset deletion flag (\lstinline|d = 0|) and then return \lstinline|ptr|.
Insertions occur bottom-up and follow the basic \citeauthor{fraser2004practical} algorithm \cite{fraser2004practical},
taking the separation of deletion flags and nodes into account.

The authors provide high level correctness and linearizability proofs as well as a model for the
SPIN model checker. Performance has been shown to compare favorably to both
\citeauthor{sundell2003fast} and \citeauthor{shavit2000skiplist} queues, with throughput improved by
up to $80\%$.

@@

\subsection{\citeauthor{wimmer2013data}} \label{sec:wimmer}
